An Adolescent Mini-Neptune in the Kepler Field

### 2021-07-30 paper/ms.tex
update ms for rotation

% gaia target sample numbers
\newcommand{\nkinematic}{3{,}298} % "fullfaint" kinematic sample, core+halo (CG18+KC19+M21). from any call of earhat.helpers._get_fullfaint_dataframes
\newcommand{\nnbhd}{13{,}843} % make_compstar_NGC_2516_sourcelist.py
\newcommand{\ncore}{1{,}106}  % "fullfaint" kinematic sample, CG18. from any call of earhat.helpers._get_fullfaint_dataframes
\newcommand{\nhalo}{2{,}192} % "fullfaint" kinematic sample, KC19+M21. from any call of earhat.helpers._get_fullfaint_dataframes

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f15a.pdf}
			\includegraphics[width=0.49\textwidth]{f15b.pdf}
		}

		\vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.8\textwidth]{f15c.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf The \cn\ is $35\pm15$\,Myr old.}  {\it Top:} HR diagram of 
    the \cn, IC\,2602 ($\approx35$\,Myr), the Pleiades ($\approx 115$\,Myr), and the field.
    The right panel provides a closer view of stars on the
    pre-main-sequence.
    The \cn\ and IC\,2602 are approximately the same isochronal age.
    {\it Bottom:} TESS and Kepler rotations periods and dereddened Gaia
    colors, with the Pleiades
    \citep[125\,Myr;][]{rebull_rotation_2016a} and Praesepe
    \citep[650\,Myr;][]{douglas_poking_2017} shown for reference.
    Stars of a given color (mass) move up through time due to magnetic
    braking.  The \cn\ has a 40 Myr gyrochronal age, as calibrated
    against the LDB-age of IC\,2602.  {\bf todo: caption improve.}.
   \label{fig:age}
	}
\end{figure*}

@@

\subsection{Rotation Periods}
%FIXME TODO REFERENCE LOWER PANELS
{\bf Jason Curtis todo}
We collected the previously measured
{\bf XXX} Kepler rotation periods (CITEP MCQUILLAN2014), and measured
{\bf YYY} new stellar rotation periods for cluster members observed by
TESS.  We also performed an isochronal analysis.  The results are
shown in Figure~\ref{fig:age}.  From the ensemble perspective, which
provides the strongest constraints (SEE CITET SODERBLOM 2014), the age
of the star is $XX \pm YY$\,Myr.  Individual measurements of the
star's rotation period and lithium abundance corroborate this
conclusion.

For photometric and astrometric binaries, see the appendix!

@@

Figure~\ref{fig:XYZvtang} shows stars reported by
\citet{kounkel_untangling_2019} to be in the group ``Theia 73'', which
was cross-matched by \citealt{kounkel_untangling_2019} as being
``Stephenson 1''.  For our Figure, galactic positions were calculated
and plotted only for stars with parallax signal-to-noise exceeding 20.
The location of the Sun is shown on the plots.  The smattering of
reported cluster members (gray and black points) has a significantly
different structure that the cluster initially identified by
\citet{stephenson_possible_1959} and corroborated by
\citet{eggen_photometric_1968}.  While the non-uniform ``clumps''
might be part of a real structure, they could also be an artifact of
the data processing steps performed by
\citet{kounkel_untangling_2019}.  We therefore opted to only consider
stars in the immediate kinematic and spatial vicinity of \sn.  The
tangential velocities relative to \sn\ are shown in the bottom right
panel.  These are computed by assuming that every star has the same
three-dimensional spatial velocity as \sn, where we assume a systemic
radial velocity of $-16.7 \pm 0.2$\,\kms\ based on the reconnaissance
spectra obtained by A.~Howard on HIRES and D.~Latham on TRES.  The
relevant projection effects are then taken into account, as discussed
by {\it e.g.}, \citet{Meingast2021} and \citet{bouma_2021_ngc2516}.
We performed the actual selection by then manually drawing lassos with
the interactive \texttt{glue} visualization tool
\citep{beaumont_2014_13866} in the four projections shown in
Figure~\ref{fig:XYZvtang}.  While bonafide members likely exist
outside of our selection region (and our selection also includes some
field star interlopers), our aim is to verify the existence of the
cluster in the vicinity of Kepler\,1627, and to measure its age.  The
procedure we have adopted enables both tasks.

\section{Binarity and Stellar Rotation for $\delta$\,Lyr Cluster Members}

f15supp.pdf

>>>>

% gaia target sample numbers
\newcommand{\nkinematic}{1{,}201} % deltalyr_kc19_cleansubset_withreddening.csv
% all from oneoff_calculate_rotation_numbers.py
\newcommand{\noriginal}{3{,}071} % Vizier; KC19 Theia 73 member count
\newcommand{\nwithtess}{924} % How many stars have TESS data?
\newcommand{\nkinwithtess}{391} % How many kinematically selected stars have TESS data?
\newcommand{\nkinwithtessandcrowding}{192} % How many kinematically selected stars have TESS data and meeting the crowding cutoff (nequal<=0, nclose<=1)?
\newcommand{\nkindefaultcleaning}{157} % How many of these meet the "defaultcleaning" criteria (LSP>0.1, Prot<15, and crowding cut)?
\newcommand{\nrotgood}{147} % nkindefaultcleaning, minus outliers.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f15a.pdf}
			\includegraphics[width=0.469\textwidth]{f15b.pdf}
		}

		\vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.8\textwidth]{f15c.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf The \cn\ is $35\pm15$\,Myr old.}  {\it Top:} HR diagram of 
    the \cn, IC\,2602 ($\approx35$\,Myr), the Pleiades ($\approx 115$\,Myr), and the field.
    The right panel provides a closer view of stars on the
    pre-main-sequence.
    The \cn\ and IC\,2602 are approximately the same isochronal age.
    {\it Bottom:} TESS and Kepler rotations periods and dereddened Gaia
    colors, with the Pleiades (125\,Myr)
    and Praesepe (650\,Myr)
    shown for reference
    \citep{rebull_rotation_2016a,douglas_poking_2017}.
    Kepler\,1627 is shown with a yellow star.
     {\bf todo: caption improve.}.
   \label{fig:age}
	}
\end{figure*}

@@

\subsection{Stellar Rotation Periods}

Of the \noriginal\ candidate \cn\ members reported by
\citet{kounkel_untangling_2019}, TESS data were available for
\nwithtess\ stars.  We extracted light curves using the nearest pixel
to each star, and regressed them against systematics using the causal
pixel model implemented by \citet{hattorio_2021_cpm} in the
\texttt{unpopular} package.  We then measured candidate rotation
periods using a Lomb-Scargle periodogram
\citep{lomb_1976,scargle_studies_1982,astropy_2018}.  To enable cuts
on crowding, we queried the Gaia source catalog for stars within a
$21\farcs0$ radius of the target star (1 TESS pixel).  We recorded the
number of stars within the aperture with greater brightness than the
target star, and with brightness within 1.25 and 2.5 TESS magnitudes
of the target star.  {\bf We then supplemented these measurements with XXX Kepler
rotation periods (CITEP MCQUILLAN2014)}.

We cleaned the candidate TESS rotation period measurements through a
combination of automated and manual steps.  As a matter of scope, we
restricted our attention to the \nkinwithtess\ stars in
the spatial and kinematic proximity of Kepler\,1627, as discussed in
Appendix~\ref{app:kinematicselection}.  We also only focused on
stars for which no companions were known with a brightness exceeding one-tenth of
the target star within a $21\farcs0$ radius: there were 
\nkinwithtessandcrowding\ stars that met these requirements, and that
had TESS data available.
For plotting purposes we then imposed a selection based on the strength of the
rotation period signal itself: we required the Lomb Scargle power to
exceed 0.1, and the period to be below 15\,days.

The lower panel of Figure~\ref{fig:age} shows the resulting
\nkindefaultcleaning\ stars.  The majority of these stars fall below
the ``slow sequence'' of the Pleiades, consistent with a
gyrochronological age for the \cn\ below 100\,Myr.  Approximately 10 stars 
appear as outliers above the ``slow sequence''.  Assuming that they
are all false positives ({\it i.e.}, field interlopers), the rotation
period detection fraction would be
\nrotgood/\nkinwithtessandcrowding$\approx 77\%$.

Binarity is known to be an important factor in the density of stars
falling underneath the ``slow sequence''
\citep{meibom_effect_2007,gillen_ngts_2020,bouma_2021_ngc2516}.
Appendix~\ref{app:rotationbinarity} discusses this topic, and finds
evidence for similar effects to those that have been previously
discussed in the literature.

\subsection{Age of Kepler\,1627}

%FIXME FIXME FIXME
From the ensemble perspective, which
provides the strongest constraints (SEE CITET SODERBLOM 2014), the age
of the star is $XX \pm YY$\,Myr.  Individual measurements of the
star's rotation period and lithium abundance corroborate this
conclusion.

@@

Figure~\ref{fig:XYZvtang} shows stars reported by
\citet{kounkel_untangling_2019} to be in the group ``Theia 73'', which
was cross-matched by Kounkel \& Covey as being ``Stephenson 1''.
\citet{kounkel_untangling_2019} reported \noriginal\ stars to be
present in this cluster.  For our Figure, galactic positions were
calculated and plotted only for stars with parallax signal-to-noise
exceeding 20.  The location of the Sun is shown on the plots.  The
smattering of reported cluster members (gray and black points) has a
significantly different structure that the cluster initially
identified by \citet{stephenson_possible_1959} and corroborated by
\citet{eggen_photometric_1968}.  While the non-uniform ``clumps''
might be part of a real structure, they could also be an artifact of
the data processing steps performed by
\citet{kounkel_untangling_2019}.  We therefore opted to only consider
stars in the immediate kinematic and spatial vicinity of \sn.  The
tangential velocities relative to \sn\ are shown in the bottom right
panel.  These are computed by assuming that every star has the same
three-dimensional spatial velocity as \sn, where we assume a systemic
radial velocity of $-16.7 \pm 0.2$\,\kms\ based on the reconnaissance
spectra obtained by A.~Howard on HIRES and D.~Latham on TRES.  The
relevant projection effects are then taken into account, as discussed
by {\it e.g.}, \citet{Meingast2021} and \citet{bouma_2021_ngc2516}.
We performed the actual selection by then manually drawing lassos with
the interactive \texttt{glue} visualization tool
\citep{beaumont_2014_13866} in the four projections shown in
Figure~\ref{fig:XYZvtang}.  While bonafide members likely exist
outside of our selection region (and our selection also includes some
field star interlopers), our aim is to verify the existence of the
cluster in the vicinity of Kepler\,1627, and to measure its age.  The
procedure we have adopted enables both tasks.

\section{Binarity and Stellar Rotation for $\delta$\,Lyr Cluster Members}
\label{app:rotationbinarity}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f15supp.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Binarity indicators for rotators in \cn.} 
    Data are as in the lower panel of
    Figure~\ref{fig:age}.  Astrometric binaries ($\mathrm{RUWE}>1.2$)
    appear in red; photometric binaries ($>$0.3\,mag above an
    empirical isochrone; 20\% of stars) are orange.  
    Over $0.6<\bpmrpo<1.5$, the detected rotation periods in binaries tend to 
    fall below the slow sequence.
    %FIXME!
		% Over $0.5<\bpmrpo<1.3$, Set $\mathcal{B}$ has \replaced{281}{523} stars.
		% Dividing these stars into ``slow'' and ``fast'' sequences by eye,
		% the fraction of stars showing signs of binarity in the slow and fast
		% subsamples are \replaced{30\% (63/210)}{27\% (106/289)} and
		% \replaced{56\% (40/71)}{51\% (68/134)}, respectively.
		\label{fig:binarity}
	}
\end{figure*}

Figure~\ref{fig:binarity} shows the rotation-color diagram for \cn\
members, with the points colored according to indicators of binarity.
The binaries tend to be redder and have shorter rotation periods.  We
know of two possible explanations based on selection effects, and two
based on physics.

Possible selection effects include {\it i)} binaries have a component
that contributes additional red light to the system, which could skew
the color measurement of the primary; and {\it ii)} the unresolved
binary companions could contaminate the rotation period measurement
\citep[{\it e.g.},][Section~5.1]{stauffer_rotation_2016}.

The two possible physical effects that could be relevant are tidal
locking and pre-main-sequence disk locking.  Tidal locking has been
argued to be an unlikely explanation for the frequency of rapid
rotators due to the population statistics; a more likely scenario
would be that the presence of the binary leads to faster disk
dispersal, enabling the primary to contract to more rapid rotation
periods than possible for single stars
\citep{meibom_effect_2007,bouma_cluster_2020}.
